\bf An Unitary RG-based Auxiliary Model Approach\\ to Strongly-Correlated Systems

### 2023-03-02 notes/notes.tex
--no-message

\section{Single-particle Green's function}
In the previous part, we obtained, among other states, the ground state \(\ket{\Psi_\text{gs}}\) of the bulk Hamiltonian in terms of the ground-states \(\ket{\psi_\text{aux}}\) of the auxiliary models.
In this section, we will relate real-sapce Greens functions of the bulk lattice to those of the auxiliary model. We will assume that the auxiliary model Hilbert space has the same dimensions as that of the bulk lattice model.
We define the time-domain Greens function between two points \(\vec R\) and \(\vec R + \vec r\) as
	G(\vec r, t) = -i\theta(t) \braket{\Psi_\text{gs} | \left\{ c(\vec R + r,t), c^\dagger(\vec R) \right\} | \Psi_\text{gs}}
We also define a Green operator \(\mathcal{G}(\omega)\):
	\mathcal{G}(\omega) = \frac{1}{\omega - (H_\text{H-H} - E_\text{gs})}~,
where \(\omega\) is the frequency scale for the excitations. This operator can be related to the Greens function through the relation (eq.~\eqref{inv_G_func})
\begin{equation}\begin{aligned}\label{greens_func_bulk}
	G(\vec r, \omega) &= \braket{\Psi_\text{gs} | c(\vec R + r) \mathcal{G}(\omega) c^\dagger(\vec R) | \Psi_\text{gs}} - \braket{\Psi_\text{gs} | c^\dagger(\vec R) \mathcal{G}(-\omega) c(\vec R + r) | \Psi_\text{gs}}\\
			  &= \sum_n\left[\braket{\Psi_\text{gs} | c(\vec R + r) \lambda^-_n \frac{\left(H_\text{H-H}\right)^n}{(\omega + E_\text{gs})^{n+1}} c^\dagger(\vec R) | \Psi_\text{gs}} + \braket{\Psi_\text{gs} | c^\dagger(\vec R) \lambda^+_n \frac{\left(H_\text{H-H}\right)^n}{(\omega - E_\text{gs})^{n+1}} c(\vec R + r) | \Psi_\text{gs}}\right]\\
			  &= \sum_n\left[ \frac{\lambda^-_n}{(\omega + E_\text{gs})^{n+1}} \mathcal{U}(\vec r)^{(n)}_> + \frac{\lambda^+_n}{(\omega - E_\text{gs})^{n+1}} \mathcal{U}(\vec r)^{(n)}_<\right]
where we have defined \(\lambda_n^\pm\) using \(1/(1\pm y) = \sum_n \lambda^\pm_n y^n\), and the retarded and advanced propagation matrix elements \(\mathcal{U}(\vec r)^{(n)}_>  = \braket{\Psi_\text{gs} | c(\vec R + r) \left(H_\text{H-H}\right)^n c^\dagger(\vec R) | \Psi_\text{gs}}, \quad \mathcal{U}(\vec r)^{(n)}_<  = \braket{\Psi_\text{gs} | c^\dagger(\vec R) \left(H_\text{H-H}\right)^n c(\vec R + r) | \Psi_\text{gs}}\).
These matrix elements will now be expressed in terms of auxiliary model properties, starting with the retarded one. By combining with eq.~\eqref{siam_to_hubb} and eq.~\eqref{bloch-states}, we get
\begin{equation}\begin{aligned}\label{retarded_matrix_element}
	\mathcal{U}(\vec r)^{(n)}_> = \frac{1}{\lambda_{\vec k_0}^2 N^2}\sum_{\vec R_i, \vec R_j} e^{i\left( \vec R_j - \vec R_i \right) \cdot \vec k_0}\braket{\psi_\text{aux}(\vec R_i) | c_{\vec R + \vec r, \sigma} \left[\sum_{\vec R_c}\mathcal{H}\left( \vec R_c \right)\right]^n  c_{\vec R, \sigma}^\dagger | \psi_\text{aux}(\vec R_j)}~,
where \(\vec k_0\) is the crystal momentum in the ground-state, and \(\ket{\psi}_\text{aux}(\vec R) = \frac{1}{w}\sum_{\vec a}^{\text{NN of }\vec R} \ket{\psi}_\text{aux}(\vec R, \vec a)\). The exponentiated sum can be simplified by inserting the momentum eigenbasis \(\left\{ \ket{\vec k}\bra{\vec k}\right\} \) of the auxiliary model Hilbert space, and shifting all the Hamiltonians \(\mathcal{H}\left( \vec R_c \right)\) to the origin \(\vec R_0\), using translation operators:

@@

	G(\vec r, \Omega) = \sum_{\vec k}\left[\frac{\theta(|\vec k - \vec k_0| - |\vec k_F|) e^{i\left(\vec k - \vec k_0\right) \cdot \vec r}}{\omega + N\varepsilon_\text{gs} - N\mathcal{H}_{\vec k, \vec k}} + \frac{\theta(|\vec k_F| - |\vec k_0 - \vec k|) e^{-i\left(\vec k - \vec k_0\right) \cdot \vec r}}{\omega - N\varepsilon_\text{gs} + N\mathcal{H}_{\vec k, \vec k}}\right]

@@

	G(\vec r, \Omega) = \sum_{\vec q}\left[\frac{\theta(|\vec q| - |\vec k_F|) e^{i\vec q \cdot \vec r}}{\omega + N\varepsilon_\text{gs} - N\mathcal{E}^p_{\vec q}} + \frac{\theta(|\vec k_F| - |- \vec q|) e^{-i\vec q \cdot \vec r}}{\omega - N\varepsilon_\text{gs} + N\mathcal{E}^h_{\vec q}}\right]~,
where the matrix elements \(\mathcal{E}^p\) and \(\mathcal{E}^h\) for the particle and hole excitations are given by \(\mathcal{E}^p_{\vec q} = \mathcal{H}_{\vec k_0 + \vec q, \vec k_0 + \vec q} = \braket{\vec k_0 | c_{\vec q} \mathcal{H} c^\dagger_{\vec q} | \vec k_0}\) and \(\mathcal{E}^h_{\vec q} = \mathcal{H}_{\vec k_0 - \vec q, \vec k_0 - \vec q} = \braket{\vec k_0 | c^\dagger_{\vec q} \mathcal{H} c_{\vec q} | \vec k_0}\).

>>>>

\section{Single-particle Green's function}
In the previous part, we obtained, among other states, the ground state \(\ket{\Psi_\text{gs}}\) of the bulk Hamiltonian in terms of the ground-states \(\ket{\psi_\text{aux}}\) of the auxiliary models.
In this section, we will relate real-sapce Greens functions of the bulk lattice to those of the auxiliary model. We will assume that the auxiliary model Hilbert space has the same dimensions as that of the bulk lattice model.
We define the time-domain Greens function between two points \(\vec R\) and \(\vec R + \vec r\) as
\begin{equation}\begin{aligned}
	G(\vec r, t) = -i\theta(t) \braket{\Psi_\text{gs} | \left\{ c(\vec R + r,t), c^\dagger(\vec R) \right\} | \Psi_\text{gs}}
\end{aligned}\end{equation}
We also define a Green operator \(\mathcal{G}(\omega)\):
\begin{equation}\begin{aligned}
	\mathcal{G}(\omega) = \frac{1}{\omega - (H_\text{H-H} - E_\text{gs})}~,
\end{aligned}\end{equation}
where \(\omega\) is the frequency scale for the excitations. This operator can be related to the Greens function through the relation (eq.~\eqref{inv_G_func})
\begin{equation}\begin{aligned}\label{greens_func_bulk}
	G(\vec r, \omega) &= \braket{\Psi_\text{gs} | c(\vec R + r) \mathcal{G}(\omega) c^\dagger(\vec R) | \Psi_\text{gs}} - \braket{\Psi_\text{gs} | c^\dagger(\vec R) \mathcal{G}(-\omega) c(\vec R + r) | \Psi_\text{gs}}\\
			  &= \sum_n\left[\braket{\Psi_\text{gs} | c(\vec R + r) \lambda^-_n \frac{\left(H_\text{H-H}\right)^n}{(\omega + E_\text{gs})^{n+1}} c^\dagger(\vec R) | \Psi_\text{gs}} + \braket{\Psi_\text{gs} | c^\dagger(\vec R) \lambda^+_n \frac{\left(H_\text{H-H}\right)^n}{(\omega - E_\text{gs})^{n+1}} c(\vec R + r) | \Psi_\text{gs}}\right]\\
			  &= \sum_n\left[ \frac{\lambda^-_n}{(\omega + E_\text{gs})^{n+1}} \mathcal{U}(\vec r)^{(n)}_> + \frac{\lambda^+_n}{(\omega - E_\text{gs})^{n+1}} \mathcal{U}(\vec r)^{(n)}_<\right]
\end{aligned}\end{equation}
where we have defined \(\lambda_n^\pm\) using \(1/(1\pm y) = \sum_n \lambda^\pm_n y^n\), and the retarded and advanced propagation matrix elements \(\mathcal{U}(\vec r)^{(n)}_>  = \braket{\Psi_\text{gs} | c(\vec R + r) \left(H_\text{H-H}\right)^n c^\dagger(\vec R) | \Psi_\text{gs}}, \quad \mathcal{U}(\vec r)^{(n)}_<  = \braket{\Psi_\text{gs} | c^\dagger(\vec R) \left(H_\text{H-H}\right)^n c(\vec R + r) | \Psi_\text{gs}}\).
These matrix elements will now be expressed in terms of auxiliary model properties, starting with the retarded one. By combining with eq.~\eqref{siam_to_hubb} and eq.~\eqref{bloch-states}, we get
\begin{equation}\begin{aligned}\label{retarded_matrix_element}
	\mathcal{U}(\vec r)^{(n)}_> = \frac{1}{\lambda_{\vec k_0}^2 N^2}\sum_{\vec R_i, \vec R_j} e^{i\left( \vec R_j - \vec R_i \right) \cdot \vec k_0}\braket{\psi_\text{aux}(\vec R_i) | c_{\vec R + \vec r, \sigma} \left[\sum_{\vec R_c}\mathcal{H}\left( \vec R_c \right)\right]^n  c_{\vec R, \sigma}^\dagger | \psi_\text{aux}(\vec R_j)}~,
\end{aligned}\end{equation}
where \(\vec k_0\) is the crystal momentum in the ground-state, and \(\ket{\psi}_\text{aux}(\vec R) = \frac{1}{w}\sum_{\vec a}^{\text{NN of }\vec R} \ket{\psi}_\text{aux}(\vec R, \vec a)\). The exponentiated sum can be simplified by inserting the momentum eigenbasis \(\left\{ \ket{\vec k}\bra{\vec k}\right\} \) of the auxiliary model Hilbert space, and shifting all the Hamiltonians \(\mathcal{H}\left( \vec R_c \right)\) to the origin \(\vec R_0\), using translation operators:
\begin{equation}\begin{aligned}
	\left[\sum_{\vec R_c}\mathcal{H}\left( \vec R_c \right)\right]^n = \left[\sum_{\vec R_c}T^\dagger_{\vec R_c -\vec R_0} \mathcal{H}\left( \vec R_0 \right)T_{\vec R_c -\vec R_0} \right]^n = \sum_{\left\{ \vec R_{i} \right\} }T^\dagger_{\vec R_1 -\vec R_0} \mathcal{H}\left( \vec R_0 \right)T^\dagger_{\vec R_2 -\vec R_1}\ldots \mathcal{H}\left( \vec R_0 \right)T_{\vec R_n -\vec R_{0}}
\end{aligned}\end{equation}

@@

	G(\vec r, \omega) = \sum_{\vec k}\left[\frac{\theta(|\vec k - \vec k_0| - |\vec k_F|) e^{i\left(\vec k - \vec k_0\right) \cdot \vec r}}{\omega + N\varepsilon_\text{gs} - N\mathcal{H}_{\vec k, \vec k}} + \frac{\theta(|\vec k_F| - |\vec k_0 - \vec k|) e^{-i\left(\vec k - \vec k_0\right) \cdot \vec r}}{\omega - N\varepsilon_\text{gs} + N\mathcal{H}_{\vec k, \vec k}}\right]

@@

	G(\vec r, \omega) = \sum_{\vec q}\left[\frac{\theta(|\vec q| - |\vec k_F|) e^{i\vec q \cdot \vec r}}{\omega + N\varepsilon_\text{gs} - N\mathcal{E}^p_{\vec q}} + \frac{\theta(|\vec k_F| - |- \vec q|) e^{-i\vec q \cdot \vec r}}{\omega - N\varepsilon_\text{gs} + N\mathcal{E}^h_{\vec q}}\right]~,
where the matrix elements \(\mathcal{E}^p\) and \(\mathcal{E}^h\) for the particle and hole excitations are given by \(\mathcal{E}^p_{\vec q} = \mathcal{H}_{\vec k_0 + \vec q, \vec k_0 + \vec q} = \braket{\vec k_0 | c_{\vec q} \mathcal{H} c^\dagger_{\vec q} | \vec k_0}\) and \(\mathcal{E}^h_{\vec q} = \mathcal{H}_{\vec k_0 - \vec q, \vec k_0 - \vec q} = \braket{\vec k_0 | c^\dagger_{\vec q} \mathcal{H} c_{\vec q} | \vec k_0}\). For models with strong local correlations, it might be more convenient to express these \(k-\)space single-particle matrix elements in terms of real-space fluctuations:
\mathcal{E}^h_{\vec q} = \sum_{\vec x_{+},\vec x_{-}}e^{i\left( \vec x_+ - \vec x_- \right) \cdot \vec q}\braket{\vec k_0 | c^\dagger_{\vec x_-} \mathcal{H} c_{\vec x_+} | \vec k_0}, \quad \mathcal{E}^p_{\vec q} = \sum_{\vec x_{+},\vec x_{-}}e^{i\left(\vec x_+ - \vec x_- \right) \cdot \vec q}\braket{\vec k_0 | c_{\vec x_+} \mathcal{H} c^\dagger_{\vec x_-} | \vec k_0}
Guided by the fact that our auxiliary model has an emergent self-consistency at the impurity site, we retain distances \(|\vec x_+ - \vec x_-|\) up to one lattice spacing \(\vec a\) from the impurity, in the summation:
	\mathcal{E}^h_{\vec q} \to \braket{\vec k_0 | c^\dagger_{\vec R_0} \mathcal{H}\left(\vec R_0\right) c_{\vec R_0} | \vec k_0} + \sum_{\vec a}\left(e^{i\vec a \cdot \vec q}\braket{\vec k_0 | c^\dagger_{\vec R_0 + \vec a} \mathcal{H}\left(\vec R_0\right) c_{\vec R_0}| \vec k_0} + \text{h.c.}\right),\\
	\quad \mathcal{E}^p_{\vec q} \to \braket{\vec k_0 | c_{\vec R_0} \mathcal{H}\left(\vec R_0\right) c^\dagger_{\vec R_0} | \vec k_0} + \sum_{\vec a}\left(e^{i\vec a \cdot \vec q}\braket{\vec k_0 | c_{\vec R_0} \mathcal{H}\left(\vec R_0\right) c_{\vec R_0 + \vec a}^\dagger| \vec k_0} + \text{h.c.}\right)

\clearpage
